The Quench-Action method in integrable spin chains: A Monte Carlo approach

### 2015-09-14 draft/fs_qa_v1.tex
rev 14/09/2015

The string centers $\lambda_{n;\gamma}$ are obtained by solving the so-called 
Bethe-Gaudin-Takahashi equations
%
\begin{equation}
\label{bgt-eq}
2L\theta_n(\lambda_{n;\gamma})=2\pi I_{n;\gamma}+\sum\limits_{(m,
\beta)\ne(n,\gamma)}\Theta_{m,n}(\lambda_{n;\gamma}-\lambda_{m;\beta}).  
\end{equation}
%
The generalized scattering phases $\Theta_{m,n}$ read 
%
\begin{eqnarray}
\nonumber\fl\Theta_{m,n}(x)\equiv\left\{\begin{array}{cc}
\theta_{|n-m|}(x)+\!\!\!\!\!\sum
\limits_{r=1}^{(n+m-|n-m|-1)/2}\!\!\!\!\!2\theta_{|n-m|+2r}(x)
+\theta_{n+m}(x) & \quad\mbox{if}
\quad n\ne m\\\fl\sum\limits_{r=1}^{n-1}2\theta_{2r}(x)+
\theta_{2n}(x) & \quad\mbox{if}\quad n=m
\end{array}\right.
\end{eqnarray}
%
with $\theta_\alpha(x)\equiv 2\arctan(x/\alpha)$. Here $I_{n;\gamma}$ are the 
Bethe-Takahashi quantum numbers associated with $\lambda_{n;\gamma}$. 
The solutions of~\eref{bgt-eq}, and the Bethe states thereof, are naturally 
classified according to their ``string content'' ${\mathcal S}\equiv\{s_n
\}_{n=1}^M$, with $s_n$ the number of $n$-strings. Clearly, the constraint 
$\sum_{n=1}^{M}n s_n=M$ has to be satisfied. It can be shown that 
the BGT quantum numbers $I_{n;\gamma}$ associated with the $n$-strings are 
integers and half-integers for $L-s_n$ odd and even, respectively. 
An upper bound for the BGT quantum numbers can be derived as~\cite{taka-book} 
%
\begin{equation}
|I_{n;\gamma}|\le I^{(MAX)}_{n}\equiv\frac{1}{2}(L-1-\sum
\limits_{m=1}^Mt_{m,n}s_m),
\label{bt_qn_bound}
\end{equation}
%
where $t_{m,n}\equiv 2\mbox{min}(n,m)-\delta_{m,n}$. Using the string hypothesis 
Bethe states energy eigenvalue~\eref{ba-ener} becomes
%
\begin{equation}
\label{ener-str}
E=-\sum_{n,\gamma}\frac{2n}{\lambda_{n;\gamma}^2+n^2}. 
\end{equation}
%

@@

Here we consider the overlap formula for the Neel state~\eref{neel-ov} in the 
limit $L\to\infty$, assuming that the rapidities form perfect strings, i.e., 
$\delta_{n;\gamma}^j=0$ in~\eref{str-hyp}.  
Then it is possible to rewrite~\eref{neel-ov} in terms of the string centers 
$\tilde\lambda_{n;\alpha}$ only. Here we restrict ourselves to rapidity 
configurations with no zero-momentum strings. Our results are not valid for 
zero-rapidity strings.

It is convenient to first split the indices $i,j$ of $G^\pm_{ij}$ as $i=(n,\gamma,i)$ 
and $j=(m,\gamma',j)$, with $n,m$ being the length of the strings, $\gamma,\gamma'$ 
labelling the corresponding string centers, and $i,j$ the components of the 
two strings. 
Using~\eref{G-pm} and~\eref{K}, one has that for two consecutive rapidities in the same 
string, i.e., for $m=n,\gamma=\gamma',|i-j|=2$, the matrices $G^{\pm}_{jk}$ become 
ill-defined in the thermodynamic limit. Precisely, since $K_{1}(\tilde\lambda_{n;\gamma}^i-
\tilde\lambda_{n;\gamma}^{i+1})\sim 1/(\delta_{n;\gamma}^i-\delta_{n;\gamma}^{i+1})$, 
$G_{ij}^\pm$ diverges in the thermodynamic limit. Importantly, as the same type of 
divergence occur in both $G^+$ and $G^-$, their ratio (cf.~\eref{neel-ov}), and 
the overlaps, are finite.

The result reads 
%
\begin{equation}
\fl \frac{1}{2}G^+_{(n,\alpha)(m,\beta)}=\left\{\begin{array}{cc}
L\theta_n'(\lambda_{n;\alpha}) -\sum\limits_{(\ell,\gamma)\ne(n,\alpha)}\Big[\Theta'_{n,\ell}
(\lambda_{n;\alpha}-\lambda_{\ell;\gamma}) & \quad\textrm{if}\,(n,\alpha)=(m,\beta) \\
+\Theta'_{n,\ell}(\lambda_{n;\alpha}+\lambda_{\ell;\gamma})\Big] & \\ \\
\Theta'_{n,m}
(\lambda_{n;\alpha}-\lambda_{m;\beta})+\Theta'_{n,m}
(\lambda_{n;\alpha}+\lambda_{m;\beta}) & \quad\textrm{if}\,(n,\alpha)\ne(m,\beta)
\end{array}\right.
\end{equation}
%
Here $\theta_n'(x)\equiv d\theta_n(x)/dx=2n/(n^2+x^2)$ and $\Theta'(x)\equiv d\Theta(x)/dx$.

For $G^-_{ij}$ one obtains

\begin{equation}
\fl\frac{1}{2}G^-_{(n,\alpha)(m,\beta)}=\left\{\begin{array}{cc}
\fl(L-1)\theta'_n(\lambda_{n;\alpha})-2\sum\limits_{k=1}^{n-1}\theta'_k(\lambda_{n;\alpha})
& \textrm{if}\,(n,\alpha)= (m,\beta)\\
-\hspace{-.5cm}\sum\limits_{(\ell,\gamma)\ne(n,\alpha)}\Big[\Theta'_{n,\ell}
(\lambda_{n;\alpha}-\lambda_{\ell;\gamma})+\Theta'_{n,\ell}
(\lambda_{n;\alpha}+\lambda_{\ell;\gamma})\Big] \\\\
\Theta'_{n,m}
(\lambda_{n;\alpha}-\lambda_{m;\beta})-\Theta'_{n,m}
(\lambda_{n;\alpha}+\lambda_{m;\beta})) & \textrm{if}\,(n,\alpha)\ne(m,\beta)
\end{array}\right.
\end{equation}
%

These would require the knowledge of the precise form of the string 
deviations, i.e., the dependence of the string deviations on $L$, as it has 
been pointed out in Ref.~\cite{calabrese-2014}.

Finally, the multiplicative prefactor in~\eref{neel-ov} for the generic $n$-string 
can be rewritten as 
%
\begin{equation}
\prod\limits_{a=1}^n\frac{\sqrt{(\lambda^a_{n;\alpha})^2+1}}{4\lambda^a_{n;\alpha}}=
\frac{1}{4^n}\left(\frac{\sqrt{n^2+\lambda^2_{n;\alpha}}}{\lambda_{n;\alpha}}
\prod\limits_{k=0}^{\lceil n/2\rceil-1}\frac{(2k)^2+\lambda^2_{n;\alpha}}{(2k+1)^2+
\lambda^2_{n;\alpha}}\right)^{{\mathcal P}},
\end{equation}
%
with ${\mathcal P}=+$ and ${\mathcal P}=-$ for even and odd strings, respectively.

@@

In the thermodynamic limit the expression for the overlap~\eref{neel-ov} becomes 
%
\begin{equation}
\label{neel-ov-th}
\fl\lim_{L\to\infty}\left[\frac{\langle N|\{\pm\lambda_j\}_{j=1}^m,n_\infty
\rangle}{|||\{\lambda_j\}_{j=1}^m,n_\infty\rangle||}\right]=\frac{L}{2}\Big(
n_\infty\log(2)+\sum_{n=1}^\infty\int_0^\infty d\lambda\rho_n(\lambda)[g_n(\lambda)
+2n\log(4)]\Big), 
\end{equation}
%
with 
%
\begin{eqnarray}
g_n(\lambda)=\sum_{l=1}^{n-1}\Big[f_{n-1-2l}(\lambda)-f_{n-2l}(\lambda)\Big],\\
f_n(\lambda)=\log\Big(\lambda^2+\frac{n^2}{4}\Big),\\
n_\infty=1-2\sum_{m=1}^\infty m\int_{-\infty}^\infty d\lambda\rho_m(\lambda). 
\end{eqnarray}
%
Note that~\eref{neel-ov-th} is extensive, subextensive contributions 
are neglected. In particular, Eq.~\eref{neel-ov-th} that the ratio of 
determinants $\det_m(G^+)/\det_m(G^-)$ does not contribute.

@@

The Majumdar-Ghosh state $|MG\rangle$ is defined as 
%
\begin{equation}
|MG\rangle\equiv \Big(\frac{|\uparrow\downarrow\rangle-|\downarrow\uparrow\rangle}
{\sqrt{2}}\Big)^{\otimes L/2}
\end{equation}
%
The overlap between the generic eigenstate of the $XXX$ chain and the Majumdar-Ghosh 
state is obtained fro the overlap with the Neel state as 
%
\begin{equation}
\label{mg-ov}
\langle MG|\{\lambda_j\}_{j=1}^M\rangle=\prod\limits_{j=1}^M\frac{1}{\sqrt{2}}
\Big(1-\frac{\lambda_j-i}{\lambda_j+i}\Big)\langle N|\{\lambda_j\}_{j=1}^M\rangle
\end{equation}
%
The mutliplicative factor in~\eref{ov-MG}, using the string hypothesis for the generic 
$n$-string is rewritten as 
%
\begin{equation}
\prod\limits_{j=1}^M\frac{1}{\sqrt{2}}
\Big(1-\frac{\lambda_j-i}{\lambda_j+i}\Big)=2^n\prod\limits_{k=0}^{\lfloor 
n/2\rfloor}\frac{1}{[\lambda_{n;\gamma}^2+(2k+(1-(-1)^n)/2)^2]^2}
\end{equation}
%

>>>>

The string centers $\lambda_{n;\gamma}$ are obtained by solving the so-called 
Bethe-Gaudin-Takahashi equations
%
\begin{equation}
\label{bgt-eq}
2L\theta_n(\lambda_{n;\gamma})=2\pi I_{n;\gamma}+\sum\limits_{(m,
\beta)\ne(n,\gamma)}\Theta_{m,n}(\lambda_{n;\gamma}-\lambda_{m;\beta}).  
\end{equation}
%
The generalized scattering phases $\Theta_{m,n}$ read 
%
\begin{eqnarray}
\label{Theta}
\nonumber\fl\Theta_{m,n}(x)\equiv\left\{\begin{array}{cc}
\theta_{|n-m|}(x)+\!\!\!\!\!\sum
\limits_{r=1}^{(n+m-|n-m|-1)/2}\!\!\!\!\!2\theta_{|n-m|+2r}(x)
+\theta_{n+m}(x) & \quad\mbox{if}
\quad n\ne m\\\fl\sum\limits_{r=1}^{n-1}2\theta_{2r}(x)+
\theta_{2n}(x) & \quad\mbox{if}\quad n=m
\end{array}\right.
\end{eqnarray}
%
with $\theta_\alpha(x)\equiv 2\arctan(x/\alpha)$. Here $I_{n;\gamma}$ are the 
Bethe-Takahashi quantum numbers associated with $\lambda_{n;\gamma}$. 
The solutions of~\eref{bgt-eq}, and the Bethe states thereof, are naturally 
classified according to their ``string content'' ${\mathcal S}\equiv\{s_n
\}_{n=1}^M$, with $s_n$ the number of $n$-strings. Clearly, the constraint 
$\sum_{n=1}^{M}n s_n=M$ has to be satisfied. It can be shown that 
the BGT quantum numbers $I_{n;\gamma}$ associated with the $n$-strings are 
integers and half-integers for $L-s_n$ odd and even, respectively. 
An upper bound for the BGT quantum numbers can be derived as~\cite{taka-book} 
%
\begin{equation}
|I_{n;\gamma}|\le I^{(MAX)}_{n}\equiv\frac{1}{2}(L-1-\sum
\limits_{m=1}^Mt_{m,n}s_m),
\label{bt_qn_bound}
\end{equation}
%
where $t_{m,n}\equiv 2\mbox{min}(n,m)-\delta_{m,n}$. Using the string hypothesis 
Bethe states energy eigenvalue~\eref{ba-ener} becomes
%
\begin{equation}
\label{ener-str}
E=-\sum_{n,\gamma}\frac{2n}{\lambda_{n;\gamma}^2+n^2}. 
\end{equation}
%

@@

Here we consider the overlap formula for the Neel state~\eref{neel-ov} in the 
limit $L\to\infty$, assuming that the rapidities form perfect strings, i.e., 
$\delta_{n;\gamma}^j=0$ in~\eref{str-hyp}.  
Then it is possible to rewrite~\eref{neel-ov} in terms of the string centers 
$\tilde\lambda_{n;\alpha}$ only. Here we restrict ourselves to rapidity 
configurations with no zero-momentum strings, i.e., with finite string 
centers. Our results are not valid for zero-momentum strings.

First, we restrict ourselves to parity-invariant string configurations, 
i.e., considering only strings having positve string centers. We denote the 
generic parity-invariant string configuration as $\{\tilde\lambda_{n;
\gamma}\}$, where $\gamma$ labels the different non-zero string centers, and 
$n$ is the string length.

It is convenient to first split the indices $i,j$ of $G^\pm_{ij}$ as $i=(n,\gamma,i)$ 
and $j=(m,\gamma',j)$, with $n,m$ being the length of the strings, $\gamma,\gamma'$ 
labelling the corresponding string centers, and $i,j$ the components of the 
two strings.

Using~\eref{G-pm} and~\eref{K}, one has that for two consecutive rapidities in the same 
string, i.e., for $m=n,\gamma=\gamma',|i-j|=2$, the matrices $G^{\pm}_{jk}$ become 
ill-defined in the thermodynamic limit. Precisely, since $K_{1}(\tilde\lambda_{n;\gamma}^i-
\tilde\lambda_{n;\gamma}^{i+1})\sim 1/(\delta_{n;\gamma}^i-\delta_{n;\gamma}^{i+1})$, 
$G_{ij}^\pm$ diverges in the thermodynamic limit. Importantly, as the same type of 
divergence occur in both $G^+$ and $G^-$, their ratio (cf.~\eref{neel-ov}), and 
the overlaps, are finite.

The finite part of the overlaps~\eref{neel-ov} can be calculated using the same strategy 
as in Ref.~\cite{calabrese-2007,calabrese-2007-a} (see also Ref.~\cite{brockmann-2014}). 
The resulting matrix $G^+$ depends only on the ``string center'' indices $(n,\gamma)$ 
and $(m,\gamma')$ and it is given as  
%
\begin{equation}
\fl \frac{1}{2}G^+_{(n,\gamma)(m,\gamma')}=\left\{\begin{array}{cc}
L\theta_n'(\tilde\lambda_{n;\gamma}) -\sum\limits_{(\ell,\alpha)\ne(n,\gamma)}\Big[\Theta'_{n,\ell}
(\tilde\lambda_{n;\gamma}-\tilde\lambda_{\ell;\alpha}) & \quad\textrm{if}\,(n,\gamma)=(m,\gamma') \\
+\Theta'_{n,\ell}(\tilde\lambda_{n;\gamma}+\tilde\lambda_{\ell;\alpha})\Big] & \\ \\
\Theta'_{n,m}
(\tilde\lambda_{n;\gamma}-\tilde\lambda_{m;\gamma'})+\Theta'_{n,m}
(\tilde\lambda_{n;\gamma}+\tilde\lambda_{m;\gamma'}) & \quad\textrm{if}\,(n,\gamma)\ne(m,\gamma')
\end{array}\right.
\end{equation}
%
Here $\theta_n'(x)\equiv d\theta_n(x)/dx=2n/(n^2+x^2)$ and $\Theta'(x)\equiv d\Theta(x)/dx$, 
with $\Theta(x)$ as defined in~\eref{Theta}. 
Similarly, for $G^-$ one obtains 
%
\begin{equation}
\fl\frac{1}{2}G^-_{(n,\gamma)(m,\gamma')}=\left\{\begin{array}{cc}
\fl(L-1)\theta'_n(\tilde\lambda_{n;\gamma})-2\sum\limits_{k=1}^{n-1}\theta'_k(
\tilde\lambda_{n;\gamma})
& \textrm{if}\,(n,\gamma)= (m,\gamma')\\
-\hspace{-.5cm}\sum\limits_{(\ell,\alpha)\ne(n,\gamma)}\Big[\Theta'_{n,\ell}
(\tilde\lambda_{n;\gamma}-\tilde\lambda_{\ell;\alpha})+\Theta'_{n,\ell}
(\tilde\lambda_{n;\gamma}+\tilde\lambda_{\ell;\alpha})\Big] \\\\
\Theta'_{n,m}
(\tilde\lambda_{n;\gamma}-\tilde\lambda_{m;\gamma'})-\Theta'_{n,m}
(\tilde\lambda_{n;\gamma'}+\tilde\lambda_{m;\gamma'})) & \textrm{if}\,(n,\gamma)\ne(m,\gamma')
\end{array}\right.
\end{equation}
%
Note that in presence of zero-momentum strings, additional divergences as $1/(\delta_{n;
\gamma}^{i}+\delta_{n;\gamma}^{i+1})$ appear due to the term $K_1(\lambda+\mu)$ 
(cf.~\eref{K}). It turns out that the treatment of these diverges is a challenging 
task because it requires the precise knowledge of the string deviations, meaning 
their dependence on $L$, for each different type of string. Some results have been 
provided for small strings in Ref.~\cite{calabrese-2014}.

Finally, using the string hypothesis and the parity-invariance condition, the multiplicative 
prefactor in~\eref{neel-ov} can be simplified. Here we focus on a generic eigenstate of 
the $XXX$ identified by $m$ pairs of finite rapidities. 
Note that due to parity invariance and the exclusion of zero-momentum strings, only 
strings of length up to $m$ are allowed. The ``reduced'' string content 
identifying the eigenstate is denoted as $\widetilde{\mathcal S}=\{\tilde s_1,\dots,\tilde s_{m}\}$, 
with $\tilde s_n$ the number of parity-invariant pairs of $n$-strings. Using the string 
hypothesis and~\eref{neel-ov}, one can write 
%
\begin{equation}
\fl\quad\prod\limits_{j=1}^m\frac{\sqrt{\tilde\lambda_j^2+1}}{4\tilde\lambda_j}=
\frac{1}{4^m}\prod\limits_{j=1}^m\prod\limits_{\ell=1}^{\tilde s_j}\left[\frac{\sqrt{j^2+
\lambda^2_{j;\ell}}}{\lambda_{j;\ell}}
\prod\limits_{k=0}^{\lceil j/2\rceil-1}\frac{(2k)^2+\lambda^2_{j;\ell}}{(2k+1)^2+
\lambda^2_{j;\ell}}\right]^{(-1)^j}.
\end{equation}
%

@@

In the thermodynamic limit $L\to\infty$ the expression for the extensive part of the 
overlap~\eref{neel-ov} can be written as~\cite{brockmann-2014} 
%
\begin{equation}
\label{neel-ov-th}
\fl-\lim_{L\to\infty}\log\left[\frac{\langle N|\{\pm\lambda_j\}_{j=1}^m,n_\infty
\rangle}{|||\{\lambda_j\}_{j=1}^m,n_\infty\rangle||}\right]=\frac{L}{2}\Big(
n_\infty\log2+\sum_{n=1}^\infty\int_0^\infty d\lambda\rho_n(\lambda)[g_n(\lambda)
+2n\log(4)]\Big), 
\end{equation}
%
where  
%
\begin{equation}
\fl\quad g_n(\lambda)=\sum_{l=1}^{n-1}\Big[f_{n-1-2l}(\lambda)-f_{n-2l}(\lambda)\Big],\quad
\textrm{with}\quad f_n(\lambda)=\log\Big(\lambda^2+\frac{n^2}{4}\Big),
\end{equation}
and
\begin{equation}
n_\infty=1-2\sum_{m=1}^\infty m\int_{-\infty}^\infty d\lambda\rho_m(\lambda). 
\end{equation}
%
Note that~\eref{neel-ov-th} is extensive. Specifically, in Eq.~\eref{neel-ov-th} 
subleading (i.e., subextensive) contributions originating from the determinant
ratio $\det_m(G^+)/\det_m(G^-)$ are neglected.

@@

The Majumdar-Ghosh state $|MG\rangle$ is defined as 
%
\begin{equation}
|MG\rangle\equiv \Big(\frac{|\uparrow\downarrow\rangle-|\downarrow\uparrow\rangle}
{\sqrt{2}}\Big)^{\otimes L/2}. 
\end{equation}
%
The overlap between a generic eigenstate of the $XXX$ chain and the Majumdar-Ghosh 
state can be obtained from the N\'eel state overlap in Eq.~\eref{neel-ov} 
as~\cite{pozsgay-2014} 
%
\begin{equation}
\label{mg-ov}
\langle MG|\{\pm\lambda_j\}_{j=1}^m\rangle=\prod\limits_{j=1}^m\frac{1}{2}
\Big(1-\frac{\lambda_j-i}{\lambda_j+i}\Big)
\Big(1+\frac{\lambda_j+i}{\lambda_j-i}\Big)
\langle N|\{\pm\lambda_j\}_{j=1}^m\rangle
\end{equation}
%
The mutliplicative factor in~\eref{ov-MG}, using the string hypothesis for the generic 
$n$-string is rewritten as 
%
\begin{equation}
\prod\limits_{j=1}^M\frac{1}{\sqrt{2}}
\Big(1-\frac{\lambda_j-i}{\lambda_j+i}\Big)=2^n\prod\limits_{k=0}^{\lfloor 
n/2\rfloor}\frac{1}{[\lambda_{n;\gamma}^2+(2k+(1-(-1)^n)/2)^2]^2}
\end{equation}
%
